A survey of generic IND-CCA2 transformations

### 2024-07-09 faster-fo-transform/cca-transform-survey/main.tex
Need to write out the implicit rejection + rigid PKE

\begin{figure}[H]
    \center
    \begin{minipage}{0.3\textwidth}
        \begin{algorithm}[H]
            \caption{$\keygen^{\not\bot}$}

        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.29\textwidth}

    \end{minipage}
    \hfill
    \begin{minipage}{0.3\textwidth}

    \end{minipage}
    \caption{$U^{\not\bot}$ routines}
\end{figure}

>>>>

\begin{figure}[H]
    \center
    \begin{minipage}{0.33\textwidth}
        \begin{algorithm}[H]
            \caption{$\keygen^{\not\bot}$}\label{alg:u-notbot-keygen}
            \begin{algorithmic}[1]
                \State $(\pk, \sk^\prime) \leftsample \keygen(1^\lambda)$
                \State $s \leftsample \mathcal{M}_\pke$
                \State $\sk = (\sk^\prime, s)$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.33\textwidth}
        \begin{algorithm}[H]
            \caption{$\encap^{\not\bot}(\pk)$}\label{alg:u-notbot-encap}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}_\pke$
                \State $c \leftsample \encrypt(\pk, m)$
                \State $K \leftarrow H(m, c)$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \begin{algorithm}[H]
            \caption{$\decap^{\not\bot}(\sk, c)$}\label{alg:u-notbot-decap}
            \begin{algorithmic}[1]
                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
                \If{$\hat{m} \in \mathcal{M}_\pke$}
                    \State $K \leftarrow H(\hat{m}, c)$
                    \State \Return $K$
                \EndIf
                \State \Return $H(s, c)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{$U^{\not\bot}$ routines}
\end{figure}

The security of $\kem^{\not\bot}$, similar to the security of $\kem^\bot$, derives from the indistinguishability between pseudorandom $K \leftarrow H(m, c)$ and truly random $K \leftsample \mathcal{K}_\kem$ under the random oracle model, unless the $\kem$ adversary somehow queries $H$ on $(m^\ast, c^\ast)$, in which case an adversary against the underlying $\pke$ with access to a $\pco$ can detect this special query use it to win the \monospace{OW-PCA} game. In addition, because $\decap^{\not\bot}$ will always return random-looking values regardless of the validity of the input ciphertext, the decapsulation oracle can be simulated without needing a ciphertext validation oracle. The $\kem$ adversary will be able to distinguish between a true decapsulation oracle and a simulated decapsulation oracle, but only by querying $H$ on $(s, c)$ for some $c$. Since $s$ is uniformly random, each query has a $\norm{\mathcal{M}_\pke}^{-1}$ chance of hitting $s$, so across $q_H$ hash queries to $H$, the probability of hitting $s$ at least once is at most $\frac{q_H}{\norm{\mathcal{M}_\pke}}$.

\begin{theorem}
    For every \monospace{IND-CCA} adversary $B$ against $\kem^{\not\bot}$ that makes $q_H$ hash queries to $H$, there exists an \monospace{OW-PCA} adversary $A$ against the underyling $\pke$ such that:

    \begin{equation*}
        \monospace{Adv}(B) \leq \frac{q_H}{\norm{\mathcal{M}_\pke}} + \monospace{Adv}(A)
    \end{equation*}
\end{theorem}

\subsubsection{KEM with explicit rejection from rigid PKE}
If the input $\pke$ is rigid (definition \ref{def:rigidity}), then the $\pco$ can be simulated with indistinguishability from the true $\pco$ except for when decryption failure occurs. Given $\pke = (\keygen, \encrypt, \decrypt)$ and hash function $H: \mathcal{M}_\pke \rightarrow \mathcal{K}_\kem$, the $U^\bot_m$ transformation outputs a $\kem^\bot_m = (\keygen, \encap^\bot_m, \decap^\bot_m)$ where the key generation routine remains unchanged. The encapsulation and decapsulation routines are described in figure \ref{fig:u-bot-m-routines}

\begin{figure}[H]
    \center
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\encap^\bot_m(\pk)$}\label{alg:encap-bot-m}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}_\pke$
                \State $c \leftarrow \encrypt(\pk, m)$
                \State $K \leftarrow H(m)$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\decap^\bot_m(\sk, c)$}\label{alg:decap-bot-m}
            \begin{algorithmic}[1]
                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
                \If{$\hat{m} \in \mathcal{M}_\pke$}
                    \State \Return $H(m)$
                \EndIf
                \State \Return $\bot$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{$\kem^\bot_m$ routines}\label{fig:u-bot-m-routines}
\end{figure}

Under the random oracle model, no adversary $B$ can distinguish between the pseudorandom $K_0 \leftarrow H(m)$ and the truly random $K_1 \leftsample \mathcal{K}_\kem$ unless $B$ queries $H$ on $m^\ast$.  Because rigidity means $\encrypt(\pk, m^\ast) = c^\ast$ is equivalent to $\decrypt(\sk, c^\ast) = m^\ast$, if $H$ is queried on $m^\ast$, a second adversary $A$ against the $\pke$ can detect the special query and use it to win the \monospace{OW-VA} game.

\begin{theorem}
    For every \monospace{IND-CCA2} adversary $B$ against $\kem^\bot_m$, there exists an \monospace{OW-VA} adversary $A$ against the underlying $\pke$ such that

    \begin{equation*}
        \monospace{Adv}(B) \leq \monospace{Adv}(A) + \delta \cdot (q_H + q_D)
    \end{equation*}
\end{theorem}

\subsubsection{KEM with implicit rejection from rigid PKE}
% TODO: write out this section
